---
title: "Case 01 EC"
author: "Alice Jiang"
date: "2/17/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(qwraps2)
options(qwraps2_markup = "markdown")
library(dplyr)
library(knitr)
library(tidyr)
library(survminer)
library(survival)
```

```{r data, include=FALSE}
dig <- read.csv("dig.csv")
dig$TRTMT <- as.factor(dig$TRTMT)
```

### Table 1
Note: Digoxin = 1 and Placebo = 0. NAs have been removed in certain calculations when needed, resulting in some proportion fractions with different denominator from the total no. of patients in the group. 
```{r table1, echo=FALSE, results="asis"}
# https://cran.r-project.org/web/packages/qwraps2/vignettes/summary-statistics.html
summary1 <-
  list(
    "Characteristics" =
      list(
        "Age (yr) - mean (sd)" = ~ qwraps2::mean_sd(AGE, digits = 1),
        "Ejection fraction - mean (sd)" = ~ qwraps2::mean_sd(EJF_PER, digits = 1),
        "Median duration of CHF - mo" = ~ median(CHFDUR, digits = 1, na.rm = TRUE)
      ),
    "% of patients" =
      list(
        "Female sex" = ~ qwraps2::n_perc(SEX == 2, digits = 1),
        "Nonwhite race" = ~ qwraps2::n_perc(RACE == 2, digits = 1),
        "Age >70 yr" = ~ qwraps2::n_perc(AGE > 70, digits = 1),
        "Cardiothoracic ratio >0.55" = ~ qwraps2::n_perc(CHESTX > 0.55, digits = 1, na_rm = TRUE),
        "Previous digoxin use" = ~ qwraps2::n_perc(DIGUSE == 1, digits = 1)
      ),
    "Method of assessing ejection fraction" =
      #1=Radionuclide, 2=Angiography, 3=2-D Echo
      list(
        "Radionuclide ventriculography" = ~ qwraps2::n_perc(EJFMETH == 1, digits = 1),
        "Two-dimensional echocardiography" = ~ qwraps2::n_perc(EJFMETH == 3, digits = 1),
        "Contrast angiography" = ~ qwraps2::n_perc(EJFMETH == 2, digits = 1)
      ),
    "NYHA class" =
      list(
        "I" = ~ qwraps2::n_perc(FUNCTCLS == 1, digits = 1, na_rm = TRUE),
        "II" = ~ qwraps2::n_perc(FUNCTCLS == 2, digits = 1, na_rm = TRUE),
        "III" = ~ qwraps2::n_perc(FUNCTCLS == 3, digits = 1, na_rm = TRUE),
        "IV" = ~ qwraps2::n_perc(FUNCTCLS == 4, digits = 1, na_rm = TRUE)
      ),
    "No. of signs or symptoms of CHF" =
      list(
        "0" = ~ qwraps2::n_perc(NSYM == 0, digits = 1),
        "1" = ~ qwraps2::n_perc(NSYM == 1, digits = 1),
        "2" = ~ qwraps2::n_perc(NSYM == 2, digits = 1),
        "3" = ~ qwraps2::n_perc(NSYM == 3, digits = 1),
        ">=4" = ~ qwraps2::n_perc(NSYM > 3, digits = 1)
      ),
    "Medical history" =
      list(
        "Previous myocardial infarction" = ~ qwraps2::n_perc(PREVMI == 1, 
                                                             digits = 1, na_rm = TRUE),
        "Current angina" = ~ qwraps2::n_perc(ANGINA == 1, digits = 1, na_rm = TRUE),
        "Diabetes" = ~ qwraps2::n_perc(DIABETES == 1, digits = 1),
        "Hypertension" = ~ qwraps2::n_perc(HYPERTEN == 1, digits = 1, na_rm = TRUE)
      ),
    "Concomitant medications" =
      list(
        "Diuretics" = ~ qwraps2::n_perc(DIURET == 1 |
                                          DIURETK == 1, digits = 1, na_rm = TRUE),
        "ACE inhibitors" = ~ qwraps2::n_perc(ACEINHIB == 1, digits = 1),
        "Nitrates" = ~ qwraps2::n_perc(NITRATES == 1, digits = 1),
        "Other vasodilators" = ~ qwraps2::n_perc(VASOD == 1, digits = 1, na_rm = TRUE)
      ),
    "Daily dose of study medication prescribed" =
      list(
        "0.125mg" = ~ qwraps2::n_perc(DIGDOSE == 0.125, digits = 1, na_rm = TRUE),
        "0.250mg" = ~ qwraps2::n_perc(DIGDOSE == 0.25, digits = 1, na_rm = TRUE),
        "0.375mg" = ~ qwraps2::n_perc(DIGDOSE == 0.375, digits = 1, na_rm = TRUE),
        "0.500mg" = ~ qwraps2::n_perc(DIGDOSE == 0.5, digits = 1, na_rm = TRUE)
      ),
    "Primary cause of CHF" =
      #1=Ischemic, 2=Hypertensive, 3=Valvular,
      #4=Idiopathic, 5=Alcohol related, 6=Other
      list(
        "Ischemic" = ~ qwraps2::n_perc(CHFETIOL == 1, digits = 1, na_rm = TRUE),
        "Nonischemic" = ~ qwraps2::n_perc(CHFETIOL != 1, digits = 1, na_rm = TRUE)
      ),
    "Primary cause of CHF - Nonischemic" =
      list(
        "Idiopathic" = ~ qwraps2::n_perc(CHFETIOL == 4, digits = 1, na_rm = TRUE),
        "Hypertensive" = ~ qwraps2::n_perc(CHFETIOL == 2, digits = 1, na_rm = TRUE),
        "Other" = ~ qwraps2::n_perc(
          CHFETIOL == 6 | CHFETIOL == 5 | CHFETIOL == 3,
          digits = 1,
          na_rm = TRUE
        )
      )
  )
summary_table(dplyr::group_by(dig, TRTMT), summary1)
```

### Commentary

Overall, I have been able to reproduce the exact same results as the NEJM paper. There were only a few discrepancies where the results are off by a few decimal place. 

In Table 1, the standard deviation of ejection fraction in the digoxin group is off by 0.1 (8.8 instead of 8.9). The proportion of patients in the digoxin group that are nonwhite (14.3% vs. 14.4%), or show >= 4 signs of CHF (80.2% vs. 80.1%), or had nitrates as concomitant medication (42.2% vs. 42.1%), or had other nonischemic cause of CHF (5.5% vs. 5.4%) are all off by 0.1%. The proportion of patients in the placebo group who had 0.250mg daily dose of study medication prescribed (70.1% vs. 70.0%), or had idiopathic (nonischemic) cause of CHF (14.2% vs. 14.1%) are also off by 0.1%. Lastly, for the proportion of patients with ischemic cause of CHF, the placebo group is off by 0.3% (70.7% vs. 70.4%) and the digoxin group is off by 0.2% (71% vs. 70.8%). 

In Table 4, the no. of randomized patients in digoxin group with CT ratio <=0.55 is off by 1 count (2221 vs. 2220) and the risk ratio for NYHA class (III or IV) is off by 0.01 (0.69 vs. 0.70).

These discrepancies could potentially be due to difference in my treatment of missing data and that of the paper. 


### Table 4
```{r, include=FALSE}
dig <- dig %>%
  mutate(treatment = case_when(
    TRTMT == 1 ~ "Digoxin", 
    TRTMT == 0 ~ "Placebo"
  )) %>%
  mutate(ejection_fraction = case_when(
    EJF_PER < 25 ~ "<0.25",
    TRUE ~ "0.25-0.45") # EJF_PER (25-45)
  ) %>%
  mutate(prior_digoxin_use = case_when(
    DIGUSE == 1 ~ "Yes",
    DIGUSE == 0 ~ "No"
  )) %>%
  mutate(cause = case_when(
    CHFETIOL == 1 ~ "ischemic", 
    !is.na(CHFETIOL) & CHFETIOL != 1 ~ "nonischemic") 
  ) %>%
  mutate(ct_ratio = case_when(
    CHESTX > 0.55 ~ ">0.55",
    TRUE ~ "<=0.55") # CT ratio <= 0.55
  ) %>%
  mutate(NYHA = case_when(
    FUNCTCLS == 1 | FUNCTCLS == 2 ~ "I or II",
    FUNCTCLS == 3 | FUNCTCLS == 4 ~ "III or IV")
  )
```

```{r, include=FALSE}
# combined outcome of death/hospitalization due to worsening heart failure
subdig <- dig %>%
  filter(DWHF == 1) #WHF == 1 | REASON == 1
```

```{r absolute_difference_%, include=FALSE}
abs_diff_ci <- function(p1, p2, n1, n2) {
  diff <- p1 - p2
  z <- 1.96 # 95%
  se <- sqrt(p1 * (1-p1) / n1 + p2 * (1-p2) / n2)
  ll <- diff - z * se
  ul <- diff + z * se
  print(paste0("Absolute Difference and 95% CI in %: ", 
               format(round(diff*100, digits = 1), nsmall = 1), " (", 
               format(round(ll*100, digits = 1), nsmall = 1), " to ", 
               format(round(ul*100, digits = 1), nsmall = 1), ")"))
}
```


#### Ejection fraction
```{r ejection_fraction, message = FALSE}
# no. of patients with >=1 event
subdig %>%
  group_by(ejection_fraction, treatment) %>%
  summarise(n = n()) %>%
  spread(treatment, n) %>%
  kable()

# no. randomized
dig %>%
  group_by(ejection_fraction, treatment) %>%
  summarise(n = n()) %>%
  spread(treatment, n) %>%
  kable()

# 0.25-0.45
prop_dig1 <- 613 / 2270
prop_placebo1 <- 735 / 2273
abs_diff_ci(prop_dig1, prop_placebo1, 2270, 2273)

# <0.25
prop_dig2 <- 428 / 1127
prop_placebo2 <- 556 / 1130
abs_diff_ci(prop_dig2, prop_placebo2, 1127, 1130)

# risk ratio
## 0.25-0.45
temp1 <- dig %>%
  filter(ejection_fraction == "0.25-0.45") 
coxph(Surv(DWHFDAYS, DWHF) ~ TRTMT, data = temp1) %>%
  gtsummary::tbl_regression(exp = TRUE)

## <0.25
temp2 <- dig %>%
  filter(ejection_fraction == "<0.25") 
coxph(Surv(DWHFDAYS, DWHF) ~ TRTMT, data = temp2) %>%
  gtsummary::tbl_regression(exp = TRUE)
```

#### Previous use of digoxin
```{r prev_use_digoxin, message=FALSE}
# no. of patients with >=1 event
subdig %>%
  group_by(treatment, prior_digoxin_use) %>%
  summarise(n = n()) %>%
  spread(treatment, n) %>%
  kable()

# no. randomized
dig %>%
  group_by(treatment, prior_digoxin_use) %>%
  summarise(n = n()) %>%
  spread(treatment, n) %>%
  kable()

# Yes previous use of digoxin
prop_dig1 <- 550 / 1498
prop_placebo1 <- 688 / 1519
abs_diff_ci(prop_dig1, prop_placebo1, 1498, 1519)

# No previous use of digoxin
prop_dig2 <- 491 / 1899
prop_placebo2 <- 603 / 1884
abs_diff_ci(prop_dig2, prop_placebo2, 1899, 1884)

# risk ratio
## Yes
temp1 <- dig %>%
  filter(DIGUSE == 1) 
coxph(Surv(DWHFDAYS, DWHF) ~ TRTMT, data = temp1) %>%
  gtsummary::tbl_regression(exp = TRUE)

## No
temp2 <- dig %>%
  filter(DIGUSE == 0) 
coxph(Surv(DWHFDAYS, DWHF) ~ TRTMT, data = temp2) %>%
  gtsummary::tbl_regression(exp = TRUE)
```

#### Cause of heart failure
```{r ischemic_nonischemic, message=FALSE}
# no. of patients with >=1 event
subdig %>%
  filter(!is.na(cause)) %>%
  group_by(treatment, cause) %>%
  summarise(n = n()) %>%
  spread(treatment, n) %>%
  kable()

# no. randomized
dig %>%
  filter(!is.na(cause)) %>%
  group_by(treatment, cause) %>%
  summarise(n = n()) %>%
  spread(treatment, n) %>%
  kable()

# ischemic cause of heart failure 
prop_dig1 <- 731 / 2405
prop_placebo1 <- 873 / 2398
abs_diff_ci(prop_dig1, prop_placebo1, 2405, 2398)

# nonischemic cause of heart failure 
prop_dig2 <- 306 / 983
prop_placebo2 <- 413 / 996
abs_diff_ci(prop_dig2, prop_placebo2, 983, 996)

# risk ratio
## ischemic
temp1 <- dig %>%
  filter(cause == "ischemic") 
coxph(Surv(DWHFDAYS, DWHF) ~ TRTMT, data = temp1) %>%
  gtsummary::tbl_regression(exp = TRUE)

## nonischemic
temp2 <- dig %>%
  filter(cause == "nonischemic") 
coxph(Surv(DWHFDAYS, DWHF) ~ TRTMT, data = temp2) %>%
  gtsummary::tbl_regression(exp = TRUE)
```

#### Cardiothoracic ratio
```{r CT_ratio, message=FALSE}
# no. of patients with >=1 event
subdig %>%
  group_by(treatment, ct_ratio) %>%
  summarise(n = n()) %>%
  spread(treatment, n) %>%
  kable()

# no. randomized
dig %>%
  group_by(treatment, ct_ratio) %>%
  summarise(n = n()) %>%
  spread(treatment, n) %>%
  kable()

# cardiothoracic ratio <=0.55
prop_dig1 <- 600 / 2221
prop_placebo1 <- 724 / 2233
abs_diff_ci(prop_dig1, prop_placebo1, 2221, 2233)

# cardiothoracic ratio >0.55
prop_dig2 <- 441 / 1176
prop_placebo2 <- 567 / 1170
abs_diff_ci(prop_dig2, prop_placebo2, 1176, 1170)

# risk ratio
## <=0.55
temp1 <- dig %>%
  filter(ct_ratio == "<=0.55") 
coxph(Surv(DWHFDAYS, DWHF) ~ TRTMT, data = temp1) %>%
  gtsummary::tbl_regression(exp = TRUE)

## >0.55
temp2 <- dig %>%
  filter(ct_ratio == ">0.55") 
coxph(Surv(DWHFDAYS, DWHF) ~ TRTMT, data = temp2) %>%
  gtsummary::tbl_regression(exp = TRUE)
```

#### NYHA class
```{r NYHA, message=FALSE}
# no. of patients with >=1 event
subdig %>%
  filter(!is.na(NYHA)) %>%
  group_by(treatment, NYHA) %>%
  summarise(n = n()) %>%
  spread(treatment, n) %>%
  kable()

# no. randomized
dig %>%
  filter(!is.na(NYHA)) %>%
  group_by(treatment, NYHA) %>%
  summarise(n = n()) %>%
  spread(treatment, n) %>%
  kable()

# NYHA class (I or II)
prop_dig1 <- 601 / 2275
prop_placebo1 <- 739 / 2296
abs_diff_ci(prop_dig1, prop_placebo1, 2275, 2296)

# NYHA class (III or IV)
prop_dig2 <- 438 / 1118
prop_placebo2 <- 552 / 1105
abs_diff_ci(prop_dig2, prop_placebo2, 1118, 1105)

# risk ratio
## I or II
temp1 <- dig %>%
  filter(NYHA == "I or II") 
coxph(Surv(DWHFDAYS, DWHF) ~ TRTMT, data = temp1) %>%
  gtsummary::tbl_regression(exp = TRUE)

## III or IV
temp2 <- dig %>%
  filter(NYHA == "III or IV") 
coxph(Surv(DWHFDAYS, DWHF) ~ TRTMT, data = temp2) %>%
  gtsummary::tbl_regression(exp = TRUE)
```

#### Overall study population
```{r overall_pop, message=FALSE}
# no. of patients hospitalized/dead due to worsening heart failure
subdig %>%
  group_by(treatment) %>%
  summarise(n = n()) %>%
  kable()

sub_ndig <- 1041
sub_nplacebo <- 1291 

# overall study population
dig %>%
  group_by(treatment) %>%
  summarise(n = n()) %>%
  kable()

ndig <- 3397
nplacebo <- 3403

prop_dig <- sub_ndig / ndig
prop_placebo <- sub_nplacebo / nplacebo
abs_diff_ci(prop_dig, prop_placebo, ndig, nplacebo)

# risk ratio
## overall study population
coxph(Surv(DWHFDAYS, DWHF) ~ TRTMT, data = dig) %>%
  gtsummary::tbl_regression(exp = TRUE)
```